BOLD Parameter Estimation Using Particle Filters

### 2010-07-22 doc/thesis/resultsB.tex
added parameter map and parameter hist

Finally voxel 7 is a completely ambiguous time series. While the thresholds applied here
are empirical, this voxel is on the borderline of the thresholds for both mutual information
and the residual. The signal itself is extremely noisy, it should the voxel should be rejected.
However this case is reminiscent of the pure-noise tests performed in the previous chapter. 
This time series an extremely good example of the danger of false positives. In spite of the
fact that the signal oscillates far faster than the BOLD estimate, because of that, it is somehow
able to maintaint a mutual information value of $.1052$ and a residual of just $1.09992$.
As such, a clear method of detecting these sorts of regions is necessary if the results
of the particle filter algorithm are toe be trusted.

\section{Discussion}
From the maps generated, the similarities to SPM8's results are encouraging. At the 
very least the output of the particle filter seems to meet the quality of SPM. The normalization
of the residual is certainly necessary. Although there is no hard threshold on the 
normalized residual, it would appear that the normalization is providing a reasonable
ordering of regression quality. Mutual information also shows promising results.
Many of the differences between SPM and the particle filter seem to be driven
more by the pre-processing methods than the regression methd. The reason for 
SPM applying such extrensive smoothing is to combat false positives. While this is
extremely important, regions such as \autoref{comp6} show the problem with doing 
so. In all, the particle filter method does very good job of estimating parameters that
fit the measurements; and at the very least this method is a workable, albiet
computationally intensive, alternative to SPM.

>>>>

Finally voxel 7 is a completely ambiguous time series. While the thresholds applied here
are empirical, this voxel is on the borderline of the thresholds for both mutual information
and the residual. The signal itself is extremely noisy, it should the voxel should be rejected.
However this case is reminiscent of the pure-noise tests performed in the previous chapter. 
This time series an extremely good example of the danger of false positives. In spite of the
fact that the signal oscillates far faster than the BOLD estimate, because of that, it is somehow
able to maintain a mutual information value of $.1052$ and a residual of just $1.09992$.
As such, a clear method of detecting these sorts of regions is necessary if the results
of the particle filter algorithm are toe be trusted.

\section{Parameter Estimates}
Although the parameters are not uniquely identifiable by a single time-series, that
does not mean estimating them is not without benefit. The parameters still contain useful
information about the system. Additionally, as an aggregate they form a distribution of
the feasible parameters parameters for a particular patient. Therefore \autoref{fig:parammaps}
contains the parameter maps for the system and \autoref{} is a histogram across all voxels
for which the mutual information was greater than $.15$. As before, the threshold is not
scientifically derived, yet in tests this threshold provided a decent balance to remove most
of the questionably active voxels in the system.

\begin{figure}
\centering
\subfigure[$\tau_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0000_6}}
\subfigure[$\tau_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0000_7}}
\subfigure[$\alpha$, Slice 6]{\includegraphics[width=7cm]{images/param_0001_6}}
\subfigure[$\alpha$, Slice 7]{\includegraphics[width=7cm]{images/param_0001_7}}
\subfigure[$E_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0002_6}}
\subfigure[$E_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0002_7}}
\subfigure[$V_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0003_6}}
\subfigure[$V_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0003_7}}
\end{figure}

\begin{figure}
\centering
\subfigure[$\tau_f$, Slice 6]{\includegraphics[width=7cm]{images/param_0004_6}}
\subfigure[$\tau_f$, Slice 7]{\includegraphics[width=7cm]{images/param_0004_7}}
\subfigure[$\tau_s$, Slice 6]{\includegraphics[width=7cm]{images/param_0005_6}}
\subfigure[$\tau_s$, Slice 7]{\includegraphics[width=7cm]{images/param_0005_7}}
\subfigure[$\epsilon$, Slice 6]{\includegraphics[width=7cm]{images/param_0006_6}}
\subfigure[$\epsilon$, Slice 7]{\includegraphics[width=7cm]{images/param_0006_7}}
\caption{Parameter Map of two regions of interest ($M.I. > .15$).}
\label{fig:ParameterMap}
\end{figure}

Regions with poor fit cannot have reliable parameter estimates because the input did not meaningfully
correlate input with output. Therefore before creating the maps,
each parameters map was masked to regions with mutual information greater than $.15$. I chose mutual
information over the residual because in the maps comparing regression fitness, the mutual information
maps had more coherency and less randomness. Additionally, in the single voxel tests (\autoref{sec:SingleVoxelReview}) 
there was more separation between the no signal case and the signal case than there was in the 
residual metric.

\begin{figure}
\centering
\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=16cm]{images/realhist}
\caption{Histogram of parameters in active regions ($M.I. > .15$).}
\label{fig:FinalHist}
\end{figure}

The parameter maps show consistency parameter across active regions, however, 
the histogram of parameter estimates across all regions with $M.I. > .15$ is the
more interesting result (\autoref{fig:FinalHist}).

\section{Discussion}
From the maps generated, the similarities to SPM8's results are encouraging. At the 
very least the output of the particle filter seems to meet the quality of SPM. The normalization
of the residual is certainly necessary. Although there is no hard threshold on the 
normalized residual, it would appear that the normalization is providing a reasonable
ordering of regression quality. Mutual information also shows promising results.
Many of the differences between SPM and the particle filter seem to be driven
more by the pre-processing methods than the regression methd. The reason for 
SPM applying such extrensive smoothing is to combat false positives. While this is
extremely important, regions such as \autoref{fig:comp6} show the problem with doing 
so. In all, the particle filter method does very good job of estimating parameters that
fit the measurements; and at the very least this method is a workable, albiet
computationally intensive, alternative to SPM.
